---
title: "Pydpiper Test Notebook"
author: "Nick Wang"
date: '2018-08-20'
output:
  rmarkdown::github_document
params:
  start: lubridate::ymd("2018-08-20")
  pipeline: "MBM"
  run_pipeline: TRUE
  download: FALSE
  data_directory: FALSE
---
This RMarkdown notebook serves as a test bed for pydpiper's standard MBM.py or twolevel_model_building.py run. To run this Rmd file as a script on the command line:

`Rscript -e "rmarkdown::render('Pydpiper_test.Rmd', params = list(pipeline='MBM'), output_dir = 'MBM', knit_root_dir = 'MBM')"`

or 

`Rscript -e "rmarkdown::render('Pydpiper_test.Rmd', params = list(pipeline='twolevel_model_building'), output_dir = 'twolevel_model_building', knit_root_dir = 'twolevel_model_building')"`

This testbed uses functions from tidyverse, RMINC, and visualization tools in MRIcrotome.
```{r setup}
library(tidyverse)
library(RMINC)
library(grid)
library(MRIcrotome)
library(glue)

#This is a fix for knittr
knitr::opts_chunk$set(cache.lazy = FALSE)
```

Download the required data from the web. 
```{r get test-data, eval=params$download}
URL <- "repo.mouseimaging.ca/repo/Pydpiper_test_files/"
tarball <- "test-data_20180925.tar.gz"
directory <- "test-data_20180925"

if(!file.exists(tarball)) {
  file.path(URL,tarball) %>%
  download.file(destfile = tarball,
                method = 'libcurl')
  if (! directory %in% dir()) {
    tarball %>% untar(compressed="gzip")
  }
}
data_directory <- directory
```

```{r point to test-data, eval=params$data_directory}
data_directory <- params$data_directory
```

## Running pydpiper on the command line
MBM.py runs a 6 parameter alignment towards an initial model assuming the input files have a random orientation scanned in different coils/spaces (lsq6-large-rotations), then runs a 12 parameter alignment towards a linear consensus average, then a non-linear alignment towards a non-linear consensus average; MAGeT segmentation is done on the lsq6 files to segment the brains. twolevel_model_building.py essentially runs two stages of MBM.py. The first stage creates a non-linear average for each of the two groups, then the second stage creates a non-linear average for the averages of the two groups.
```{r MBM.py, eval=(params$pipeline=="MBM" & params$run_pipeline), include=params$pipeline=="MBM"}
command <- glue("MBM.py --pipeline-name=MBM_test",
"--subject-matter mousebrain",
"--num-executors 1000 --time 48:00:00",
"--csv-file {file.path(data_directory,'input.csv')}",
"--lsq6-large-rotations-tmp-dir=/tmp",
"--init-model {file.path(data_directory, 'Pydpiper-40-micron-basket-dec-2014/basket_mouse_brain_40micron.mnc')}",
"--run-maget",
"--maget-registration-method minctracc",
"--maget-atlas-library {file.path(data_directory, 'ex-vivo')}",
"--maget-nlin-protocol {file.path(data_directory, 'default_nlin_MAGeT_minctracc_prot.csv')}",
"--maget-masking-nlin-protocol {file.path(data_directory, 'default_nlin_MAGeT_minctracc_prot.csv')}",
"--lsq12-protocol {file.path(data_directory, 'Pydpiper_testing_default_lsq12.csv')}",
.sep=" ")
command %>% system
```

```{r twolevel_model_building.py, eval=(params$pipeline=="twolevel_model_building" & params$run_pipeline), include=params$pipeline=="twolevel_model_building"}
command <-  glue("twolevel_model_building.py --pipeline-name=twolevel_test",
"--subject-matter mousebrain",
"--num-executors 1000 --time 48:00:00",
"--csv-file {file.path(data_directory,'input.csv')}",
"--lsq6-large-rotations-tmp-dir=/tmp",
"--init-model {file.path(data_directory, 'Pydpiper-40-micron-basket-dec-2014/basket_mouse_brain_40micron.mnc')}",
"--run-maget",
"--maget-registration-method minctracc",
"--maget-atlas-library {file.path(data_directory, 'ex-vivo')}",
"--maget-nlin-protocol {file.path(data_directory, 'default_nlin_MAGeT_minctracc_prot.csv')}",
"--maget-masking-nlin-protocol {file.path(data_directory, 'default_nlin_MAGeT_minctracc_prot.csv')}",
"--lsq12-protocol {file.path(data_directory, 'Pydpiper_testing_default_lsq12.csv')}",
.sep=" ")
command %>% system
```

In total, there are 16 brains imaged twice, 2 weeks apart. The second set of images had volume changes artificially induced in the following regions:
```{r table, echo=FALSE}
changes <- tibble::tribble(
  ~region, ~inducedChange,
  'olfactory bulbs', 1.05,
  'striatum', 0.85,
  'cerebral cortex: occipital lobe', 0.93,
  'dentate gyrus of hippocampus', 1.10,
  'cerebellar cortex', 1.07
)
changes
```

## R Analysis
Load the consensus average and corresponding masks as `mincArray`s.
```{r consensus MBM.py, eval=params$pipeline=="MBM", include=params$pipeline=="MBM"}
consensusVol <- file.path("MBM_test_nlin", "MBM_test-nlin-3.mnc") %>%
  mincGetVolume() %>%
  mincArray
d <- consensusVol %>% dim()

consensusMaskPath <- file.path('MBM_test_nlin', 'MBM_test-nlin-3_mask.mnc') 
consensusMask <- consensusMaskPath %>% mincGetVolume
```

```{r consensus twolevel_model_building.py, eval=params$pipeline=="twolevel_model_building", include=params$pipeline=="twolevel_model_building"}
consensusVol <- file.path("twolevel_test_second_level_nlin", "twolevel_test_second_level-nlin-2.mnc") %>%
  mincGetVolume() %>%
  mincArray
d <- consensusVol %>% dim()

consensusMaskPath <- file.path('twolevel_test_second_level_nlin', 'twolevel_test_second_level-nlin-2_mask.mnc') 
consensusMask <- consensusMaskPath %>% mincGetVolume
```

Load MBM.py's useful output csv file pointing to the processed images.
```{r gfs MBM.py, eval=params$pipeline=="MBM", include=params$pipeline=="MBM"}
gfs <- file.path("analysis.csv") %>% 
  read_csv() %>% 
  filter(fwhm == 0.2) %>% 
  mutate(name = str_c(type, coil, sep="_"),
         type = fct_relevel(type, "wt")) %>%
  select(-fwhm)
```

```{r gfs twolevel_model_building.py, eval=params$pipeline=="twolevel_model_building", include=params$pipeline=="twolevel_model_building"}
gfs <- file.path("analysis.csv") %>% 
  read_csv() %>% 
  filter(fwhm == 0.2) %>% 
  mutate(name = str_c(type, coil, sep="_"),
         type = fct_relevel(type, "wt"),
         log_full_det=overall_log_full_det)%>%
  select(-fwhm)
```

Do an `anatGetAll` call on the MAGeT segmented label files. anatMatrix is a matrix of volumes, one for each region for each brain.
```{r anatMatrix}
defPath <- "test-data_20180925/ex-vivo/Dorr_2008_mapping_of_labels.csv"

anatMatrix <- anatGetAll(gfs$label_file, method="labels", defs=defPath) %>%
  anatCombineStructures(defs=defPath) %>% 
  unclass() %>%
  as_tibble()
```

Find the mean volumes for each region for the two groups.
```{r fractions}
fractions <- anatMatrix %>% 
  bind_cols(select(gfs,type)) %>% 
  gather(key=region, value=size, -type) %>% 
  group_by(type, region) %>% 
  summarise(mean_size=mean(size)) %>%
  spread(key=type, value=mean_size) %>%
  #filter(wt > 1) %>%
  mutate(fraction = mut/wt) %>% 
  arrange(desc(fraction)) %>%
  left_join(changes)
fractions
```

Also look at the regions with greatest shrinkage.
```{r bottom}
top_n(fractions, -10, fraction)
```

Are these changes significant? Call `anatLm` to fit a linear model on each brain region, and `anatFDR` to correct for multiple comparisons. Did modified regions yield significant results?
```{r anatomicalAbs}
anatMatrix <- anatMatrix %>% select(fractions$region)

avs <- anatLm(~type, gfs, anatMatrix)
qavs <- anatFDR(avs)
fractions$qvalue_abs <- qavs[,"qvalue-tvalue-typemut"]
fractions %>% 
  arrange(qvalue_abs) %>% 
  filter(!is.na(inducedChange))
```

Repeat the analysis for each region's volume relative to the total brain size by co-varying for brain size in the `anatLm` call. Do modified regions yield significant results when covarying for total brain size?
```{r anatomicalRel"}
gfs$brainVolumes <- anatMatrix %>% 
  rowSums()

avsrel <- anatLm(~type+brainVolumes, gfs, anatMatrix)
qavsrel <- anatFDR(avsrel)

fractions$qvalue_rel <- qavsrel[,"qvalue-tvalue-typemut"]
fractions %>% 
  arrange(qvalue_rel) %>% 
  filter(!is.na(inducedChange))
```

For the sake of visualization, let us do voxel-wise analysis.

```{r voxelwiseAbs, cache=TRUE}
vs <- mincLm(log_full_det ~ type, gfs, mask=consensusMaskPath)
vsFDR <- mincFDR(vs, mask=consensusMaskPath, method="FDR")
vsFDR %>% thresholds()
```
```{r voxelwiseRel, cache=TRUE}
vsrel <- mincLm(log_full_det ~ type+brainVolumes, gfs, mask=consensusMaskPath)
vsrelFDR <- mincFDR(vsrel, mask=consensusMaskPath, method="FDR")
vsrelFDR %>% thresholds()
```

```{r visualize}
sliceSeries(nrow=5, ncol=2, begin=100, end =300) %>%
  anatomy(consensusVol, range(consensusVol)[1], range(consensusVol)[2]) %>%
  overlay(mincArray(vs, "tvalue-typemut"), low=vsFDR %>% thresholds() %>% {.["0.05", "tvalue-typemut"]}, high=10, symmetric = TRUE) %>%
  addtitle("Absolute Volume Changes") %>%
  contourSliceIndicator(consensusVol, c(700,1400)) %>%
  legend("t-statistics") %>%
#######
  sliceSeries(nrow=5, ncol=2, begin=100, end = 300) %>%
  anatomy() %>% #reuse previous anatomy call's arguments
  overlay(mincArray(vsrel, "tvalue-typemut"), low=vsrelFDR %>% thresholds() %>% {.["0.05", "tvalue-typemut"]}, high=10, symmetric = TRUE) %>%
  addtitle("Relative Volume Changes") %>%
  contourSliceIndicator(consensusVol, c(700,1400)) %>%
  legend("t-statistics") %>%
  draw()
```